---
title: "Using Python in R - reticulate"
author: Jennifer HY Lin
date: 2025-10-19
draft: true
categories: 
    - Notes
format: html
---

This is something I've always wanted to try, that is using both R and Python in one place so here it is. 

Steps:

This notebook has been created using a virtual environment using uv, with the associated repository created using the project structure from uv too. This step has been done prior to the executions of the following steps.

1. Install R with version at least 4.0 or above, and Python with version at least 3.7 or above (ref: https://rviews.rstudio.com/2022/05/25/calling-r-from-python-with-rpy2/).

2. Install R's language server as shown below.

3. I've just figured out that I probably should've installed [reticulate package](https://rstudio.github.io/reticulate/) first actually ([GitHub repo here](https://github.com/rstudio/reticulate/)). It's basically a R interface to Python that'll be very useful.

* Install rmarkdown package so can preview .qmd docs.

* Go to Settings of VS Code -> under "User" -> Text Editor -> Files -> Associations -> add *.Rmd as "Item" and rmd as "Value"

4. Install R extension for VS Code (because I'm using VS Code but I think there is a very similar IDE called Positron created by Posit which may be easier to use for this case).

* Install shiny package

-> then I got the rmarkdown preview to work!

5. I'll have to try to show and run both R and Python code below **to be done**

```{r}
library(reticulate)
```

```{r}
# Run Python code
py_run_string("x = 10")
py_run_string("y = 5")
py_run_string("z = x + y")

# Access the Python variable in R
z <- py$z  # Correctly access the Python variable 'z' in R
print(z)
```

